\LARGE Performance Evaluation of Xenomai 3

### 2015-10-21 paper.tex
revised cyclictest

\subsubsection{Test scenario}
The goal of the experiment is to evaluate the realtime performances of mainline Linux kernel, PREEMPT\_RT, 
Xenomai 2 and Xenomai 3. Also, we want to find out whether the newly released Xenomai-3.0 has improved in realtime capability, compared to Xenomai-2.6.4, the latest stable version of Xenomai 2.

\subsubsection{Tools}
Cyclictest is a commonly-used latency benchmark. It is compatible with all the kernel setups we use (the recently Xenomai-3.0 release add cyclictest to its testsuite), therefore it's suitable for the purpose of measuring latencies for all setups.
Latency is a test program for the Xenomai test suite. It can measure latencies under 3 modes - user task, kernel task, or timer IRQ. It launches the test procedure in one of the user-specified contexts. In the user mode, the measurement task is handled by the 'latency' function in the program, which periodically waits and fetches the current time when resuming. Latency is then measured by calculating the difference of the actual time elapsed with the predefined period.

\subsubsection{Test Procedure}
Each experiment was run on the same hardware. Each was run for 20 minutes to generate approximately 1,000,000 samples. Test scripts were written to facilitate the experiment process.
\subsubsection{Test Load}
As previously mentioned, we measure latencies under IDLE and CPU-STRESSED conditions. In the CPU-STRESSED condition, we called the "stress" process to exert pressure on the CPU.
\subsection{Experiment Result}
\begin{figure}[H]
\begin{center}
\includegraphics[width=6cm]{img/cyclictest_idle.jpg}
\caption{Cyclictest with no stress.}
\label{cyclictest-idle}
\end{center}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}[H]
\begin{center}
\includegraphics[width=6cm]{img/cyclictest_stress.jpg}
\caption{Cyclictest with stress.}
\label{cyclictest-stress}
\end{center}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}[H]
\begin{center}
\includegraphics[width=6cm]{img/x2-idle.png}
\caption{Latency with no stress on Xenomai-2.}
\label{x2-latency-idle}
\end{center}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}[H]
\begin{center}
\includegraphics[width=6cm]{img/x2-cpu.png}
\caption{Latency with stress on Xenomai-2.}
\label{x2-latency-stress}
\end{center}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}[H]
\begin{center}
\includegraphics[width=6cm]{img/x3-idle.png}
\caption{Latency with no stress on Xenomai-3.}
\label{x3-latency-idle}
\end{center}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}[H]
\begin{center}
\includegraphics[width=6cm]{img/x3-cpu.png}
\caption{Latency with stress on Xenomai-3.}
\label{x3-latency-stress}
\end{center}
\end{figure} 
%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{tabular}{|l|r|r|r|}
\hline
\multirow{2}{*}{} & \multicolumn{3}{c|}{idle}  \\ \cline{2-4} 
 & \multicolumn{1}{l|}{min} & \multicolumn{1}{l|}{avg} & \multicolumn{1}{l|}{max} \\ \hline
Mainline Linux & 39 & 43 & 1046  \\ \hline
Preempt-RT & 16 & 21 & 52 \\ \hline
Xenomai 2 & 8 & 9 & 41 \\ \hline
Xenomai 3 & 7 & 7 & 35 \\ \hline
 & \multicolumn{3}{c|}{cpu-stressed} \\ \hline
Mainline Linux & 39 & 52 & 1097 \\ \hline
Preempt-RT & 18 & 25 & 62 \\ \hline
Xenomai 2 & 8 & 19 & 45 \\ \hline
Xenomai 3 & 6 & 16 & 37 \\ \hline
\end{tabular}\\
Table 2: cyclictest result on different kernel setups, in \textmu s\\
%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{tabular}{|l|c|r|r|}
\hline
\multirow{2}{*}{} & \multicolumn{3}{c|}{user} \\ \cline{2-4} 
 & min & \multicolumn{1}{c|}{avg} & \multicolumn{1}{c|}{max} \\ \hline
Xenomai 2 & 8.541 & 9.458 & 34.583 \\ \hline
Xenomai 3 & 8.043 & 8.853 & 33.023 \\ \hline
& \multicolumn{3}{c|}{kernel}      \\ \hline
Xenomai 2 & 6.965 & 7.708 & 27.821 \\ \hline
Xenomai 3 & 5.567 & 6.479 & 25.577 \\ \hline
& \multicolumn{3}{c|}{timer-irq}   \\ \hline
Xenomai 2 & 2.129 & 2.654 & 11.343 \\ \hline
Xenomai 3 & 2.586 & 3.079 & 10.031 \\ \hline
\end{tabular} \\
Table 3: latency result on different kernel setups, in \textmu s\\

From the data above, it shows that the latency performance of timer-irq in CPU-STRESSED situation is slightly worse on Xenomai-3(over 3 \textmu s) than on Xenomai-2(below 3 \textmu s). While this is not an occasional case(the same experiment was ran a couple times), we tried to look into the test code of latency to figure out the cause of this phenomenon.

@@

\bibliographystyle{unsrt}
\begin{thebibliography}{9}%use this if you have <=9 bib refs
	\bibitem {rtai}{\it https://www.rtai.org/}
	\bibitem {sel4}{\it http://ssrg.nicta.com.au/projects/seL4/}
	\bibitem {linux-rt}{\it http://git.kernel.org/cgit/linux/kernel/git/rt/}
	\bibitem {xenomai}{\it http://xenomai.org/}
	\bibitem {adeos} Karim Yaghmour. Adaptive Domain Environment for Operating Systems.\\
	    \url{http://www.opersys.com/ftp/pub/Adeos/adeos.pdf}
	\bibitem {ChameleonRTOS}{\it http://www.xenomai.org/documentation/slides/Xenomai-OSMB-2007-01.pdf}
	\bibitem {posix-1003-1c}{\it http://standards.ieee.org/findstds/interps/1003-1c-95\_int/}
	\bibitem {xenomai-solo}{\it https://www.osadl.org/Xenomai-SOLO.xenomai-solo.0.html}
	\bibitem {x3-applications}{\it http://xenomai.org/running-applications-with-xenomai-3-x/}
	\bibitem {bbb}{\it http://beagleboard.org/BLACK}
	\bibitem {am335x}{\it http://www.ti.com/lsds/ti/processors/sitara/arm\_cortex-a8/am335x/overview.page}
	\bibitem {debian}{\it http://debian.beagleboard.org/images/bone-debian-7.5-2014-05-14-2gb.img.xz}
	\bibitem {kernel}{\it https://github.com/beagleboard/linux}
	\bibitem {p-rt}{\it https://www.kernel.org/pub/linux/kernel/projects/rt/3.14/}
	\bibitem {p-ipipe}{\it http://download.gna.org/adeos/patches/v3.x/arm/ipipe-core-3.14.39-arm-9.patch}
	\bibitem {git-x2.6}{\it http://git.xenomai.org/xenomai-2.6.git/}
	\bibitem {git-x3}{\it http://git.xenomai.org/xenomai-3.git/}
	\bibitem {stress}{\it http://people.seas.harvard.edu/~apw/stress/}
\end{thebibliography}

>>>>

\subsubsection{Test scenario}
We prepared comprehensive statistics collection for the real-time performance results of mainline Linux kernel, PREEMPT\_RT, Xenomai 2 and Xenomai 3. In addition, we expect to figure out the performance benefits from the newly released Xenomai version 3.0, the first official release of Xenomai 3, comparing to Xenomai-2.6.4, the latest stable release of Xenomai 2. This new architecture of Xenomai 3 exhibits two real-time options, selected at build time. The dual kernel configuration of Xenomai 3, Cobalt, is a significant rework of the Xenomai 2 system. And, the single kernel natively over mainline Linux kernels, which can be supplemented by the PREEMPT\_RT to meet stricter response time requirements.

Since the single kernel configuration is built upon mainline kernel or PREEMPT\_RT, we do not have to measure its configuration separately because of the availability of PREEMPT\_RT in our benchmark matrix.

\subsubsection{Performance Tools}
\textit{Cyclictest}, maintained in \textit{rt-tests} \cite{rt-tests} package, is a commonly-used latency benchmark. It is compatible with all of the kernel setups we use (the recently Xenomai 3.0 release shipped with cyclictest in its testsuite), therefore it is suitable for the purpose of measuring latencies for all setups.

\textit{Latency} is a test program provided by the Xenomai test suite (available in both Xenomai 2 and 3). It can measure latencies under 3 modes: \textit{user task}, \textit{kernel task}, or \textit{timer IRQ}. It launches the test procedure in one of the user-specified contexts. In the user mode, the measurement task is handled by the 'latency' function in the program, which periodically waits and fetches the current time when resuming. Latency is then measured by calculating the difference of the actual time elapsed with the predefined period.

\subsubsection{Test Procedure}
Each experiment was run on the same Beaglebone Black hardware device. Each was run for 20 minutes to generate approximately 1,000,000 samples. Test scripts were written to facilitate the experiment process.

\subsubsection{Test Load}
As previously mentioned, we measure latencies under IDLE and CPU-STRESSED conditions. In the CPU-STRESSED condition, we invoked the "stress" process to exert pressure on the CPU.

\subsection{Experiment Result}
\begin{figure}[H]
\begin{center}
\includegraphics[width=6cm]{img/cyclictest_idle.jpg}
\caption{Cyclictest with no stress.}
\label{cyclictest-idle}
\end{center}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}[H]
\begin{center}
\includegraphics[width=6cm]{img/cyclictest_stress.jpg}
\caption{Cyclictest with stress.}
\label{cyclictest-stress}
\end{center}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}[H]
\begin{center}
\includegraphics[width=6cm]{img/x2-idle.png}
\caption{Latency with no stress on Xenomai-2.}
\label{x2-latency-idle}
\end{center}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}[H]
\begin{center}
\includegraphics[width=6cm]{img/x2-cpu.png}
\caption{Latency with stress on Xenomai-2.}
\label{x2-latency-stress}
\end{center}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}[H]
\begin{center}
\includegraphics[width=6cm]{img/x3-idle.png}
\caption{Latency with no stress on Xenomai-3.}
\label{x3-latency-idle}
\end{center}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}[H]
\begin{center}
\includegraphics[width=6cm]{img/x3-cpu.png}
\caption{Latency with stress on Xenomai-3.}
\label{x3-latency-stress}
\end{center}
\end{figure} 
%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{tabular}{|l|r|r|r|}
\hline
\multirow{2}{*}{} & \multicolumn{3}{c|}{idle}  \\ \cline{2-4} 
 & \multicolumn{1}{l|}{min} & \multicolumn{1}{l|}{avg} & \multicolumn{1}{l|}{max} \\ \hline
Mainline Linux & 39 & 43 & 1046  \\ \hline
Preempt-RT & 16 & 21 & 52 \\ \hline
Xenomai 2 & 8 & 9 & 41 \\ \hline
Xenomai 3 & 7 & 7 & 35 \\ \hline
 & \multicolumn{3}{c|}{cpu-stressed} \\ \hline
Mainline Linux & 39 & 52 & 1097 \\ \hline
Preempt-RT & 18 & 25 & 62 \\ \hline
Xenomai 2 & 8 & 19 & 45 \\ \hline
Xenomai 3 & 6 & 16 & 37 \\ \hline
\end{tabular}\\
Table 2: cyclictest result on different kernel setups, in \textmu s\\
%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{tabular}{|l|c|r|r|}
\hline
\multirow{2}{*}{} & \multicolumn{3}{c|}{user} \\ \cline{2-4} 
 & min & \multicolumn{1}{c|}{avg} & \multicolumn{1}{c|}{max} \\ \hline
Xenomai 2 & 8.541 & 9.458 & 34.583 \\ \hline
Xenomai 3 & 8.043 & 8.853 & 33.023 \\ \hline
& \multicolumn{3}{c|}{kernel}      \\ \hline
Xenomai 2 & 6.965 & 7.708 & 27.821 \\ \hline
Xenomai 3 & 5.567 & 6.479 & 25.577 \\ \hline
& \multicolumn{3}{c|}{timer-irq}   \\ \hline
Xenomai 2 & 2.129 & 2.654 & 11.343 \\ \hline
Xenomai 3 & 2.586 & 3.079 & 10.031 \\ \hline
\end{tabular} \\
Table 3: latency result on different kernel setups, in \textmu s\\

From the data above, it shows that the latency performance of timer-irq in CPU-STRESSED situation is slightly worse on Xenomai3 (over 3 \textmu s) than on Xenomai 2 (below 3 \textmu s). While this is not an occasional case (the same experiment was ran a couple times), we tried to look into the test code of latency to figure out the cause of this phenomenon.

@@

\bibliographystyle{unsrt}
\begin{thebibliography}{9}%use this if you have <=9 bib refs
	\bibitem {rtai}{\it https://www.rtai.org/}
	\bibitem {sel4}{\it http://ssrg.nicta.com.au/projects/seL4/}
	\bibitem {linux-rt}{\it http://git.kernel.org/cgit/linux/kernel/git/rt/}
	\bibitem {xenomai}{\it http://xenomai.org/}
	\bibitem {adeos} Karim Yaghmour. Adaptive Domain Environment for Operating Systems.\\
	    \url{http://www.opersys.com/ftp/pub/Adeos/adeos.pdf}
	\bibitem {ChameleonRTOS}{\it http://www.xenomai.org/documentation/slides/Xenomai-OSMB-2007-01.pdf}
	\bibitem {posix-1003-1c}{\it http://standards.ieee.org/findstds/interps/1003-1c-95\_int/}
	\bibitem {xenomai-solo}{\it https://www.osadl.org/Xenomai-SOLO.xenomai-solo.0.html}
	\bibitem {x3-applications}{\it http://xenomai.org/running-applications-with-xenomai-3-x/}
	\bibitem {bbb}{\it http://beagleboard.org/BLACK}
	\bibitem {am335x}{\it http://www.ti.com/lsds/ti/processors/sitara/arm\_cortex-a8/am335x/overview.page}
	\bibitem {debian}{\it http://debian.beagleboard.org/images/bone-debian-7.5-2014-05-14-2gb.img.xz}
	\bibitem {kernel}{\it https://github.com/beagleboard/linux}
	\bibitem {p-rt}{\it https://www.kernel.org/pub/linux/kernel/projects/rt/3.14/}
	\bibitem {p-ipipe}{\it http://download.gna.org/adeos/patches/v3.x/arm/ipipe-core-3.14.39-arm-9.patch}
	\bibitem {git-x2.6}{\it http://git.xenomai.org/xenomai-2.6.git/}
	\bibitem {git-x3}{\it http://git.xenomai.org/xenomai-3.git/}
	\bibitem {stress}{\it http://people.seas.harvard.edu/~apw/stress/}
	\bibitem {rt-tests}{\it https://rt.wiki.kernel.org/index.php/Cyclictest}
\end{thebibliography}
